PAGAI : a path sensitive static analyser

### 2012-05-19 papers/TAPAS2012/pagai.tex
intro

\usepackage{amsmath}
\usepackage{mathcomp,amsfonts,amssymb,MnSymbol}
\usepackage[authoryear]{natbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{tikz}

@@

    \thanks[imag.fr]{Email:
    \href{First Name.Last Name@imag.fr} {\texttt{\normalshape
        First Name.Last Name@imag.fr}}} 
    \thanks[grenoble-inp.fr]{Email:
    \href{Matthieu.Moy@grenoble-inp.fr} {\texttt{\normalshape
        Matthieu.Moy@grenoble-inp.fr}}} 
\begin{abstract} 
	We present design and implementation of PAGAI, a new static analyzer 
	working over the LLVM compiler infrastructure, that computes inductive
	invariants on the numerical variables of the program.
	PAGAI implements various state-of-the-art techniques combining Abstract
	Interpretation and Decision Procedures (SMT-solving), that allow to do a
	precise analysis of the various paths in a program.
	We thus can compare in time and precision the various combinations of
	techniques and abstract domains. We conducted extensive experiments for
	comparing these techniques on several benchmarks and GNU programs.
\end{abstract}
\begin{keyword}
	Static Analysis, Program Verification, Abstract Interpretation, Decision
	Procedure, Satisfiability Modulo Theories.
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{intro}
Static analysis aims at automatically computing properties on programs, such as
the possible values of their variables during execution. This allows to show for
instance that a program has no runtime error (such as arithmetic overflow,
division by zero), to compute loop invariants, etc. The main applications of
static analysis are compile-time optimisations, and the proof of safety
properties in critical systems.
Abstract Interpretation is a general framework for fully automatic static
analysis, that computes inductive invariants at various control points of an
imperative program. Linear Relation Analysis (LRA) is a direct application of
this framework, that computes an upper-approximation of the set of the reachable
states of a numerical program.

>>>>

\usepackage{amsmath}
\usepackage{mathcomp,amsfonts,amssymb}
\usepackage[authoryear]{natbib}
\usepackage[mathscr]{euscript}
\usepackage{algorithmic,algorithm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{tikz}

@@

    \thanks[imag.fr]{Email:
    \href{First Name.Last Name@imag.fr} {\texttt{\normalshape
        First Name.Last Name@imag.fr}}} 
    \thanks[grenoble-inp.fr]{Email:
    \href{Matthieu.Moy@grenoble-inp.fr} {\texttt{\normalshape
        Matthieu.Moy@grenoble-inp.fr}}} 
\begin{abstract} 
	We describe the design and the implementation of PAGAI,
        a new static analyzer working over the LLVM compiler infrastructure,
        which computes inductive invariants on the numerical variables of the
        analyzed program.

	PAGAI implements various state-of-the-art algorithms combining abstract
	interpretation and decision procedures (SMT-solving),
        focusing on distinction of paths inside the control flow graph while
        avoiding systematic exponential enumerations.
        It is parametric in the abstract domain in use, the algorithm,
        and the decision procedure.

	We compared the time and precision of various combinations of
	analysis algorithms and abstract domains, with extensive
        experiments both on personal benchmarks and widely available
        GNU programs.
\end{abstract}
\begin{keyword}
	Static Analysis, Program Verification, Abstract Interpretation, Decision
	Procedure, Satisfiability Modulo Theories.
\end{keyword}
\end{frontmatter}
\section{Introduction}\label{intro}
Sound static analysis automatically computes properties on programs, such as the possible values of their variables during execution.
Applications include:
showing that a program has no runtime error (such as arithmetic overflow, division by zero, array access out of bounds), as in e.g. the Astrée analyzer \cite{ASTREE_ESOP05};
computing invariants for use with assisted proof systems (such as the B method, Frama-C\footnote{http://www.frama-c.com/}), thereby lessening the burden on the user;
computing invariants for advanced optimization techniques in compilation (e.g. showing that two array cells are distinct, in order to allow instruction reordering between assignments to these cells).

Abstract Interpretation is a general framework for fully automatic static
analysis, that computes inductive invariants at various control points of an
imperative program. Linear Relation Analysis (LRA) is a direct application of
this framework, that computes an upper-approximation of the set of the reachable
states of a numerical program.
